### 2023-08-05 RunSims.Rmd
Updates simulattions and manuscrito

---
title: "Running simulations of MultipleContactNeutralMigrat.nlogo"
author: "Leonardo A. Saravia"
date: "8/11/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

needed_packages <- c(
    "tidyverse"
  , "lubridate"
  , "future"
  , "tictoc"
  , "nlrx")

lapply(needed_packages, function(x) { if(!require(x,character.only = TRUE)) install.packages(x)} )

theme_set(theme_bw())
source("R/functions.r")

#
# Where the NetLogo model is
# Setup for nlrx 
#
if( Sys.info()['nodename'] =="ls-pro") {
  simfolder <- "/home/leonardo/Dropbox/Projects/DynamicForestExtinction"

} else if(Sys.info()['nodename'] =="biologia2018") {                     # Server UNGS
  simfolder <- "/home/leonardo/Dropbox/Projects/DynamicForestExtinction"

}

# Unix default NetLogo installation path (adjust to your needs!):
#
netlogopath <- file.path("/home/leonardo/NetLogo")
modelpath <- file.path(simfolder, "MultipleContactNeutralMigrat.nlogo")
outpath <- file.path("Simulations")

# If not defined set the JAVA version of your local 
if(Sys.getenv("JAVA_HOME")==""){
  Sys.setenv(JAVA_HOME = "/usr/lib/jvm/java-11-openjdk-amd64")
  ## "/usr/lib/jvm/java-8-oracle"
}

nl <- nl(nlversion = "6.3",
         nlpath = netlogopath,
         modelpath = modelpath,
         jvmmem = 2048)

```

## Why do several small patches hold more species that few large patches?

## Lambda = 1-4 NoSelection = Neutral  Regular

* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 ) 
* Check simulations with Migration0-at-deforestation true or false to take into account that immigration happens at the borders and different kinds of deforestation patterns have different border size e.g. regular > randon block

```{r SLOSS_lambda14_sims_regular,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}

# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
# hps = habitat patch size
# dd = dispersal distance 
# brb = bird rate birds 
nl@experiment <- experiment(expname="NoSelection_pf06_hps3-61_dd1-3_brb1-4_regular",
                            outpath=outpath,
                            repetition=1,
                            tickmetrics="false",
                            idsetup="setup",
                            idgo="go",
                            runtime=0,
                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
                                             ),
                            constants = list("world-width" = 199,
                                             "world-height" = 199,
                                             "initial-population" = 0,
                                             "max-birds-species" = 100,
                                             "death-rate-birds" = 1,
                                             "migration-rate-birds" = 0.0001,
                                             "Video" = "false",
                                             "prob-frag" =  0.6,
                                             "Deforestation-at-time" = 200,
                                             "end-time" = 1500,
                                             "Migration0-at-deforestation" = "false",
                                             "deforestation-type" = "\"regular\"",
                                             "birds-behavior" = "\"NoSelection\"",
                                             "replacement-rate" = 0.0
                                             ))

#
# Run nseeds=10 times 
#
# Total number of jobs = nseeds * split = 20*2 
# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
#

nl@simdesign <- simdesign_ff(nl=nl,
                               nseeds=30)

# run in Paralell 
#
plan(multisession)
tic()
results <- run_nl_all(nl,split = 2)
#results <- run_nl_all(nl,split = 4)
toc()
plan(sequential)
names(results)

#
# Write the output 
#
setsim(nl, "simoutput") <- results 
write_simoutput(nl)
```

## Read Simulations

```{r readSimulations, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
# Read previous simulations
#
if( file.exists("Simulations/sloss_sim_ff.rds")) {

  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")

} else {
  sloss_sim <- tibble()
}

# Read netlogo simulations and 
# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
# Calculate cluster stats
#
# Old 10 - 20 simulations 
# mdl <- read_netlogo_simul(file.path(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_ff.csv"),skip=0) %>% 
# mdl <- read_netlogo_simul(file.path(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_20_ff.csv"),skip=0) %>%

mdl <- read_sloss_simul(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_regular_ff.csv")

sloss_sim <- bind_rows(sloss_sim,mdl)

saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")

netlogo_evaluate_patch_distr(mdl$find_cluster_sizes[1])

```

## Plots - No selection - regular pattern

```{r PlotsNoSelectionRegular, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}

# Neutral communities 
#
mdl <- sloss_sim %>% filter(deforestation_type == "regular",birds_behavior=="NoSelection")
names(mdl)
#
# Plot for neutral communities
# Calculate the power exponent 
# pexp = (1 - 2 * birds_dispersal_distance ) / (1 - birds_dispersal_distance ), median_dispersal_distance = 2^(1/pexp-1) 
#
mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(c(birds_at_deforestation:count_birds,max_patch),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) )))

names(mdl1)

mdl1 <- mdl1 %>%  mutate(across(max_patch_mean:max_patch_qh,sqrt))

#
# Block patch size vs max patch size
#
mdl1 %>%  
  ggplot(aes(x = max_patch_mean, y = habitat_patch_size, color = factor(habitat_patch_size))) + 
  geom_point(  size = 1.2) + guides(color = "none") #+

# require(plotly)
# plot_ly(data=mdl1, x=~birds_dispersal_distance, y=~habitat_patch_size, z=~calc_number_of_species, type="scatter3d", mode="markers", color=~calc_number_of_species, marker = list(size = 1))

mdl1 %>%  
  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
  scale_color_viridis_d() +
  facet_wrap(~ birth_rate_birds, ncol = 4) + 
  labs(color = "Dispersal\n distance") +
  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
  theme(
    legend.position = c(0.9,0.8)
    )

ggsave("Figures/Species_Habitat_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)

# Shannon-Diversity  
#
mdl1 %>%  
  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
  scale_color_viridis_d() +
  facet_wrap(~ birth_rate_birds, ncol = 4) + 
  labs(color = "Dispersal\n distance") +
  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
  theme(
    legend.position = c(0.9,0.8)
    )

ggsave("Figures/Shannon_Habitat_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)

#
# Number of species by max_patch size 
# 
mdl1  %>% #filter(birds_dispersal_distance %in% c(1.1,1.5,3)) %>%
  ggplot(aes(x = max_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)) ,  size = 1.2) +
  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
  scale_color_viridis_d() +
  facet_wrap(~ birth_rate_birds, ncol = 4) + 
  labs(color = "Dispersal\n distance") +
  theme_bw() + xlab("Max patch size") + ylab("Number of species") +
  theme(
    legend.position = c(0.9,0.8)
    )

ggsave("Figures/Species_MaxPatch_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)

#
# shannon_diversity by max_patch size 
# 
mdl1  %>% #filter(birds_dispersal_distance %in% c(1.1,1.5,3)) %>%
  ggplot(aes(x = max_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)) ,  size = 1.2) +
  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
  scale_color_viridis_d() +
  facet_wrap(~ birth_rate_birds, ncol = 4) + 
  labs(color = "Dispersal\n distance") +
  theme_bw() + xlab("Max patch size") + ylab("Shannon Diversity") +
  theme(
    legend.position = c(0.9,0.2)
    )

ggsave("Figures/Shannon_MaxPatch_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)

```

## Lambda = 1-4 / Random Block patterns

* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 )

```{r SLOSS_lambda14_simulations_block,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}

# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
# hps = habitat patch size
# dd = dispersal distance 
# brb = bird rate birds 
nl@experiment <- experiment(expname="NoSelection_pf06_hps3-61_dd1-3_brb1-4_block",
                            outpath=outpath,
                            repetition=1,
                            tickmetrics="false",
                            idsetup="setup",
                            idgo="go",
                            runtime=0,
                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path", "find-cluster-sizes"),
                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
                                             ),
                            constants = list("world-width" = 199,
                                             "world-height" = 199,
                                             "initial-population" = 0,
                                             "max-birds-species" = 100,
                                             "death-rate-birds" = 1,
                                             "migration-rate-birds" = 0.0001,
                                             "Video" = "false",
                                             "prob-frag" =  0.6,
                                             "Deforestation-at-time" = 200,
                                             "end-time" = 1500,
                                             "Migration0-at-deforestation" = "false",
                                             "deforestation-type" = "\"random block\"",
                                             "birds-behavior" = "\"NoSelection\"",
                                             "replacement-rate" = 0.0
                                             ))

#
# Run nseeds=10 times 
#
# Total number of jobs = nseeds * split = 30*2 
# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
#

nl@simdesign <- simdesign_ff(nl=nl,
                               nseeds=30)

# run in Paralell 
#
plan(multisession)
tic()
results <- run_nl_all(nl,split = 2)
#results <- run_nl_all(nl,split = 4)
toc()
plan(sequential)
names(results)

#
# Write the output 
#
setsim(nl, "simoutput") <- results 
write_simoutput(nl)
```

## Read Simulations No Selection Random Block Patterns

```{r readSimulationsBlock, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
# Read previous simulations
#
if( file.exists("Simulations/sloss_sim_ff.rds")) {

  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")

} else {
  sloss_sim <- tibble()
}

# Read netlogo simulations and 
# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
#
mdl <- read_sloss_simul(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_block_ff.csv")

sloss_sim <- bind_rows(sloss_sim,mdl)

saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")

```

## Plots Neutral No Selection Random Block Patterns Plots

```{r NeutralRandomBlockPlots, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}

# Neutral communities 
#
mdl <- sloss_sim %>% filter(deforestation_type == "random block", birds_behavior == "NoSelection")
names(mdl)
#
# Plot for neutral communities 
#
mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(c(birds_at_deforestation:count_birds,max_patch) ,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/

names(mdl1)
mdl1 <- mdl1 %>%  mutate(across(max_patch_mean:max_patch_qh,sqrt))

#
# Block patch size vs max patch size
#
mdl1 %>%  
  ggplot(aes(x = max_patch_mean, y = habitat_patch_size, color = factor(habitat_patch_size))) + 
  geom_point(  size = 1.2) + guides(color = "none") #+

#
# Number of species by patch block size
#
mdl1 %>%  
  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
  scale_color_viridis_d() +
  facet_wrap(~ birth_rate_birds, ncol = 4) + 
  labs(color = "Dispersal\n distance") +
  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
  theme(
    legend.position = c(0.9,0.8)
    )

ggsave("Figures/Species_Habitat_Dispersal_Lambda_block.png",width=8,height=6,units="in",dpi=600)

# Shannon-Diversity  
#
mdl1 %>%  
  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
  scale_color_viridis_d() +
  facet_wrap(~ birth_rate_birds, ncol = 4) + 
  labs(color = "Dispersal\n distance") +
  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
  theme(
    legend.position = c(0.9,0.8)
    )

ggsave("Figures/Shannon_Habitat_Dispersal_Lambda_block.png",width=8,height=6,units="in",dpi=600)

#
# Number of species by max_patch size 
# 
mdl1  %>% #filter(birds_dispersal_distance %in% c(1.1,1.5,3)) %>%
  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)) ,  size = 1.2) +
  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
  scale_color_viridis_d() +
  facet_wrap(~ birth_rate_birds, ncol = 4) + 
  labs(color = "Dispersal\n distance") +
  theme_bw() + xlab("Max patch size") + ylab("Number of species") +
  theme(
    legend.position = c(0.9,0.83)
    )

ggsave("Figures/Species_MaxPatch_Dispersal_Lambda_block.png",width=8,height=6,units="in",dpi=600)

#
# shannon_diversity by max_patch size 
# 
mdl1  %>% # filter(birds_dispersal_distance %in% c(1.1,1.5,3)) %>%
  ggplot(aes(x = max_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)) ,  size = 1.2) +
  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
  scale_color_viridis_d() +
  facet_wrap(~ birth_rate_birds, ncol = 4) + 
  labs(color = "Dispersal\n distance") +
  theme_bw() + xlab("Max patch size") + ylab("Shannon Diversity") +
  theme(
    legend.position = c(0.9,0.2)
    )

ggsave("Figures/Shannon_MaxPatch_Dispersal_Lambda_block.png",width=8,height=6,units="in",dpi=600)

```

## Simulations Lambda = 1-4 / Regular patterns - Hierarchical communities

* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 )

```{r Hierarchical_SLOSS_lambda14_simulations_regular,echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}

# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
# hps = habitat patch size
# dd = dispersal distance 
# brb = bird rate birds 
nl@experiment <- experiment(expname="Hierarchical_pf06_hps3-61_dd1-3_brb1-4_regular",
                            outpath=outpath,
                            repetition=1,
                            tickmetrics="false",
                            idsetup="setup",
                            idgo="go",
                            runtime=0,
                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
                                             ),
                            constants = list("world-width" = 199,
                                             "world-height" = 199,
                                             "initial-population" = 0,
                                             "max-birds-species" = 100,
                                             "death-rate-birds" = 1,
                                             "migration-rate-birds" = 0.0001,
                                             "Video" = "false",
                                             "prob-frag" =  0.6,
                                             "Deforestation-at-time" = 200,
                                             "end-time" = 1500,
                                             "Migration0-at-deforestation" = "false",
                                             "deforestation-type" = "\"regular\"",
                                             "birds-behavior" = "\"Hierarchical\"",
                                             "replacement-rate" = 0.3
                                             ))

#
# Run nseeds=10 times 
#
# Total number of jobs = nseeds * split = 30*2 
# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
#

nl@simdesign <- simdesign_ff(nl=nl,
                               nseeds=30)

# run in Paralell 
#
plan(multisession)
tic()
results <- run_nl_all(nl,split = 2)
#results <- run_nl_all(nl,split = 4)
toc()
plan(sequential)
names(results)

#
# Write the output 
#
setsim(nl, "simoutput") <- results 
write_simoutput(nl)
```

## Read Simulations hierarchical regular Patterns

```{r readSimulationsHierarchicalRegular, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
# Read previous simulations
#
if( file.exists("Simulations/sloss_sim_ff.rds")) {

  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")

} else {
  sloss_sim <- tibble()
}

# Read netlogo simulations and 
# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
#
mdl <- read_sloss_simul(outpath,"Hierarchical_pf06_hps3-61_dd1-3_brb1-4_regular_ff.csv")

sloss_sim <- bind_rows(sloss_sim,mdl)

saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")

```

# Plot for hierarchical communities

```{r ExploratoryPlotsHierarchical, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}

mdl <- sloss_sim %>% filter(deforestation_type == "regular", birds_behavior=="Hierarchical")
names(mdl)
#
# summarize data 
#
mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(birds_at_deforestation:count_birds,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
names(mdl1)

# require(plotly)
# plot_ly(data=mdl1, x=~birds_dispersal_distance, y=~habitat_patch_size, z=~calc_number_of_species, type="scatter3d", mode="markers", color=~calc_number_of_species, marker = list(size = 1))

mdl1 %>% 
  ggplot(aes(x=habitat_patch_size, y=calc_number_of_species_mean, color=birds_dispersal_distance)) + 
  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds))) +
  geom_ribbon(
    aes(
      ymin = calc_number_of_species_ql,
      ymax = calc_number_of_species_qh,
      fill = birds_dispersal_distance
    ),
    alpha = 0.3,
    colour = NA,
    show.legend = FALSE

  ) +
  scale_color_viridis_c(guide = FALSE) + 
  scale_fill_viridis_c() +
  facet_grid(birds_dispersal_distance ~ birth_rate_birds) + 
  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species")

mdl1 %>%  
  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
  scale_color_viridis_d() +
  facet_wrap(~ birth_rate_birds, ncol = 4) + 
  labs(color = "Dispersal\n distance") +
  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
  theme(
    legend.position = c(0.9,0.8)
    )

ggsave("Figures/Hierarchical_Species_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)

# Shannon-Diversity  
#
mdl1 %>%  
  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
  scale_color_viridis_d() +
  facet_wrap(~ birth_rate_birds, ncol = 4) + 
  labs(color = "Dispersal\n distance") +
  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
  theme(
    legend.position = c(0.9,0.8)
    )

ggsave("Figures/Hierarchical_Shannon_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)

```

## Simulations Lambda = 1-4 / Regular patterns - Birth Selection communities

* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 )

```{r Selection_SLOSS_lambda14_simulations_regular,echo=FALSE,message=FALSE,warning=FALSE,eval=TRUE}

# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
# hps = habitat patch size
# dd = dispersal distance 
# brb = bird rate birds 
nl@experiment <- experiment(expname="Selection_pf06_hps3-61_dd1-3_brb1-4_regular",
                            outpath=outpath,
                            repetition=1,
                            tickmetrics="false",
                            idsetup="setup",
                            idgo="go",
                            runtime=0,
                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
                                             ),
                            constants = list("world-width" = 199,
                                             "world-height" = 199,
                                             "initial-population" = 0,
                                             "max-birds-species" = 100,
                                             "death-rate-birds" = 1,
                                             "migration-rate-birds" = 0.0001,
                                             "Video" = "false",
                                             "prob-frag" =  0.6,
                                             "Deforestation-at-time" = 200,
                                             "end-time" = 1500,
                                             "Migration0-at-deforestation" = "false",
                                             "deforestation-type" = "\"regular\"",
                                             "birds-behavior" = "\"BirthSelection\"",
                                             "replacement-rate" = 0.3
                                             ))

#
# Run nseeds=10 times 
#
# Total number of jobs = nseeds * split = 30*2 
# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
#

nl@simdesign <- simdesign_ff(nl=nl,
                               nseeds=30)

# run in Paralell 
#
plan(multisession)
tic()
results <- run_nl_all(nl,split = 2)
#results <- run_nl_all(nl,split = 4)
toc()
plan(sequential)
names(results)

#
# Write the output 
#
setsim(nl, "simoutput") <- results 
write_simoutput(nl)
```

## Read Simulations Birth Selection  regular Patterns

```{r readSimulationsSelectionRegular, echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
# Read previous simulations
#
if( file.exists("Simulations/sloss_sim_ff.rds")) {

  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")

} else {
  sloss_sim <- tibble()
}

# Read netlogo simulations and 
# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
#
mdl <- read_sloss_simul(outpath,"Selection_pf06_hps3-61_dd1-3_brb1-4_regular_ff.csv")

sloss_sim <- bind_rows(sloss_sim,mdl)

saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")

```

## Not fragmented communities Lambda = 1.7 - 4

```{r SLOSS_notfrag_simulations, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
require(tidyverse)

source("R/functions.r")

# Not fragmented communities = habitat_patch_size = 100
#
mdl <- read_netlogo_simul("Simulations/NoHiBi_pf00_hpf3_dd1-3_lambda1.7-4.csv") %>% mutate(habitat_patch_size=100)

# New Simulations 
#
mdl <- read_netlogo_simul("Simulations/NoHiBi_pf00_hpf3_dd1-3_lambda1.1-4.csv") %>% mutate(habitat_patch_size=100)

sloss_sim <- bind_rows(sloss_sim,mdl)

# Add Migration0_at_deforestation=FALSE to previous simulations 
#
unique(sloss_sim$Migration0_at_deforestation)
sloss_sim <- sloss_sim %>% mutate( Migration0_at_deforestation = if_else(is.na(Migration0_at_deforestation), FALSE, Migration0_at_deforestation )) 
unique(sloss_sim$habitat_patch_size)
#sloss_sim <- sloss_sim %>% mutate( habitat_patch_size = if_else(habitat_patch_size==0, 100, habitat_patch_size ))

saveRDS(sloss_sim, "Simulations/sloss_sim.rds")

names(mdl)
#
# Plot for neutral communities 
# I filter birth_rate_birds 1.1 because these communities survive only by migration
#
mdl1 <- mdl %>% filter(birth_rate_birds!=1.1) %>% group_by(birds_dispersal_distance, birth_rate_birds,birds_behavior) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=birth_rate_birds)) + geom_point() + geom_line() + scale_color_viridis_d(name="lambda") + facet_wrap( ~birds_behavior )  + theme_bw() + ylab("Number of species") + xlab("Dispersal Distance") 
ggsave("Figures/Species_NoFrag_Dispersal_Lambda.png",width=8,height=6,units="in",dpi=600)

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=birth_rate_birds)) + geom_point() + scale_color_viridis_d() + facet_wrap( ~birds_behavior ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=count_birds,color=birth_rate_birds)) + geom_point() + geom_line() + scale_color_viridis_d(name="lambda") + facet_wrap( ~birds_behavior )  + theme_bw() + ylab("Number of indiv.") + xlab("Dispersal Distance") 
ggsave("Figures/Indiv_NoFrag_Dispersal_Lambda.png",width=8,height=6,units="in",dpi=600)

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>%  mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=birth_rate_birds,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()+ facet_wrap( ~birds_behavior )

```

## Add more simulations lambda = 2 with Migration

```{r SLOSS_combined_plots, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}

mdl <- read_netlogo_simul("Simulations/NoHiBi_pf06_hpf3-61_dd1-3_lambda2_Migrat.csv")

sloss_sim <- bind_rows(sloss_sim,mdl)

mdl <- read_netlogo_simul("Simulations/NoHiBi_pf06_hpf3-61_dd1-3_lambda2_NoMigrat.csv") 
sloss_sim <- bind_rows(sloss_sim,mdl)

mdl <- read_netlogo_simul("Simulations/NoHiBi_pf06_hpf3-61_dd1-3_lambda4_Migrat.csv") 
sloss_sim <- bind_rows(sloss_sim,mdl)

save.image()

#
# 200 x 200 - Started simulation 10:47
#
mdl <- read_netlogo_simul("Simulations/NoHiBi_200_pf06_hpf3-61_dd1-3_lambda2_NoMigrat.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)
sloss_sim <- sloss_sim %>% mutate( world_width=if_else(is.na(world_width),100,world_width))
sloss_sim <- sloss_sim %>% mutate( world_width=if_else(world_width==99,100,world_width))

saveRDS(sloss_sim, "Simulations/sloss_sim.rds")

```

## Plots of combined behaviors and patch 100

```{r SLOSS_combined_plots1, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
require(tidyverse)

source("R/functions.r")

#mdl1 <- sloss_sim %>% filter(Migration0_at_deforestation | (birth_rate_birds==2 & habitat_patch_size==100)) %>%  
sloss_sim %>% distinct(Migration0_at_deforestation, world_width)

#
# Count number of simulations
#
mdl1 <- sloss_sim %>% filter(Migration0_at_deforestation, world_width==100 ) %>%  
  count(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior)

mdl1 <- sloss_sim %>% filter(Migration0_at_deforestation, world_width==100 ) %>%  
  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior) %>%
  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )

#mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3 || birds_dispersal_distance == 1.5) %>%  
mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% NoMigration")

ggsave("Figures/Species_PatchSize_Dispersal_NoMigration.png",width=8,height=6,units="in",dpi=600)

mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_shannon_diversity,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Shannon diversity") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% NoMigration")
ggsave("Figures/Shannon_PatchSize_Dispersal_NoMigration.png",width=8,height=6,units="in",dpi=600)

#
# With migration all the time
#

mdl1 <- sloss_sim %>% filter(!Migration0_at_deforestation ) %>%  
  count(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior)

#
# Plot without SD
#
mdl1 <- sloss_sim %>% filter(!Migration0_at_deforestation & habitat_patch_size!=100 & birth_rate_birds == 2, world_width==100 ) %>%  
  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior) %>%
  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )

mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% Migration")

ggsave("Figures/Species_PatchSize_Dispersal_Migration.png",width=8,height=6,units="in",dpi=600)

mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_shannon_diversity,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Shannon diversity") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% Migration")
ggsave("Figures/Shannon_PatchSize_Dispersal_Migration.png",width=8,height=6,units="in",dpi=600)

#
# From Lamda 2 to lambda 4 Neutral
#

mdl1 <- sloss_sim %>% filter(!Migration0_at_deforestation & habitat_patch_size!=100 & birds_behavior == "NoSelection" ) %>%  
  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds) %>%
  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )

mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birth_rate_birds ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Neutral - Degraded 60% - Migration 0.001")
ggsave("Figures/Species_lambda2-4_PatchSize_Dispersal_NoSelection.png",width=8,height=6,units="in",dpi=600)

#
# From Lamda 2 to lambda 4 Hierarchical
#

mdl1 <- sloss_sim %>% filter(!Migration0_at_deforestation & habitat_patch_size!=100 & birds_behavior == "Hierarchical" ) %>%  
  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds) %>%
  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )

mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birth_rate_birds ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Hierarchical - Degraded 60% - Migration 0.001")
ggsave("Figures/Species_lambda2-4_PatchSize_Dispersal_Hierarchical.png",width=8,height=6,units="in",dpi=600)

```