### 2024-03-19 ms_SLOSS_title.Rmd
Changes to submit

---
title: "Why there are more species in several small patches vs few large patches:
  a multi-species modeling approach"
bibliography: sloss_bib.bib
csl: "ecology-letters.csl"

output:
  pdf_document: 
    includes:
      in_header: margins.sty
---

Leonardo A. Saravia ^1^^2^, Ulises Balza ^1^, Fernando Momo^3^^4^

1. Centro Austral de Investigaciones Científicas (CADIC-CONICET). Houssay 200 (9410), Ushuaia, Argentina.

2. Universidad Nacional de Tierra del Fuego (UNTDF), Yrigoyen 879 (9410), Ushuaia, Argentina

3. Instituto de Ciencias. Universidad Nacional de General Sarmiento. J.M. Gutierrez 1159 (1613), Los Polvorines, Argentina.

4. Departamento de Ciencias Básicas. Universidad Nacional de Luján. Rutas 5 y 7, (6700) Luján. Argentina.

## Acknowledgments

We acknowledge the financial support provided by the "Fondo para la Investigación Científica y Tecnológica FONCYT" (Grant No. PICT-2020-SERIEA-02628). Additionally, we extend our gratitude to the National University of General Sarmiento (UNGS), especially Santiago Doyle, for granting access to the computational resources utilized in this study.