### 2024-10-09 ms_SLOSS_title.Rmd
Updated title and Readme with zenodo

---
title: "Why there are more species in several small patches vs few large patches:
  a multi-species modeling approach"
bibliography: sloss_bib.bib
csl: "ecology-letters.csl"

output:
  pdf_document: 
    includes:
      in_header: margins.sty
---

Leonardo A. Saravia ^1^^2^, Ulises Balza ^1^, Fernando Momo^3^^4^

1. Centro Austral de Investigaciones Científicas (CADIC-CONICET). Houssay 200 (9410), Ushuaia, Argentina.

2. Universidad Nacional de Tierra del Fuego (UNTDF), Yrigoyen 879 (9410), Ushuaia, Argentina

3. Instituto de Ciencias. Universidad Nacional de General Sarmiento. J.M. Gutierrez 1159 (1613), Los Polvorines, Argentina.

4. Departamento de Ciencias Básicas. Universidad Nacional de Luján. Rutas 5 y 7, (6700) Luján. Argentina.

## Author Contributions

Leonardo Saravia, Ulises Balsa, and Fernando Momo conceived the project. Leonardo Saravia designed, developed, and performed the simulations. He analyzed the data in close consultation with all co-authors. Leonardo Saravia and Ulises Balsa wrote the first draft of the manuscript. All authors contributed to the manuscript’s revision and finalization.

## Acknowledgments

We acknowledge the financial support provided by the "Fondo para la Investigación Científica y Tecnológica FONCYT" (Grant No. PICT-2020-SERIEA-02628). Additionally, we extend our gratitude to the National University of General Sarmiento (UNGS), especially to Dr. Santiago Doyle, for granting access to the computational resources utilized in this study.

## Conflict Of Interest Statement

We declare no conflict of interest.

## Data Availability Statement

The NetLogo model is available for download on GitHub: https://github.com/lsaravia/DynamicForestExtinction and Zenodo https://doi.org/10.5281/zenodo.8157152. Data and R scripts can also be accessed on GitHub: https://github.com/EcoComplex/multiSLOSS and Zenodo https://doi.org/10.5281/zenodo.13891484.